Lowestcase and Uppestcase letters: Advances in Derp Learning

### 2021-03-21 trunk/lowercase/paper/lowercase.tex
Add code for tracing hand-made bitmaps into fonts, add hand-made font and write about it, tidy up paper a little

% \usepackage{code}
% \usepackage{cite}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% IPA symbols. safe turns off overrides for like \! which I still want
\usepackage[safe]{tipa}
\usepackage{chessboard}
% \usepackage{chessfs}
% \usepackage{adjustbox}

@@

The thing is: What if I'm even MORE ANGRY THAN I WAS BEFORE? There are
some standard sorts of typographic emphasis, like I can be {\bf BOLD
  ANGRY} or \textbf{\textit{\large BIG BOLD ITALIC UNDERLINE ANGRY}}
or { \Large \textbf{\textit{\uuline{COMBINE A LOT OF THESE ANGERS}}}},
each with its own nuances, depending on the cascading style sheet or
LaTeX class file. To be even more casual than lowercase, u can learn 2
write like this, and {\scriptsize shrink away} and also \sout{cross
  out ur words in shame in advance of them even being read}, but there
are few other options for de-emphasis. Plus, when I'm FEELING PRETTY
ANGRY, TOM, how do I capitalize that already-capitalized T in order to
show the proper reverence for your humanity?

@@

% shift-reduce conflict
% Shift is often used to increase conflict, not reduce it

@@

\begin{figure}[ht]
% \includegraphics[width=0.9 \linewidth]{manual}
\caption{TODO} \label{fig:manual}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=0.9 \linewidth]{training}
\caption{ Simultaneously training the two models. This example
  illustrates how a pair of letterforms \letterform{A} and
  \letterform{a} from the same font becomes four training examples.
  The pair straightforwardly generates an example
  \trainingexample{A}{a} for the \makelowercase\ queue, and an example
  \trainingexample{a}{A} for the \makeuppercase\ queue. Separately, we
  supply \letterform{a} to the \makelowercase\ model, simply to get
  the current output \letterform{\lowerlowera} (no model updates are
  performed). But this pair reversed becomes a training example
  \trainingexample{\lowerlowera}{a} for the \makeuppercase\ queue.
} \label{fig:cotraining}
\end{figure}

@@

\begin{figure}[ht]
\centering
  \includegraphics[width=0.99 \linewidth]{lastlayer}
  \caption{ The bottom-right corner of the weight matrix for the final
    layer of the network. This layer was added to the network after
    1.36 million rounds, initially as the identity matrix, and so can
    be thought of partly as a correction of the network's output prior
    to that round (though the remainder of the network continues to
    evolve). The x-axis is the output nodes, and the y-axis is the
    nodes of the previous layer. Note for example that the last 26
    columns look pretty different; these are the predictors for the 26
    letters, which occur in the output after the SDF pixels. Green
    means positive and red means negative, so if you are looking at
    this in a black-and-white printout, that may explain your current
    confusion. The exact diagonal is a strong green, close to 1.0,
    although over time these weights do diverge from the identity
    somewhat. In the bottom-right corner of size $26^2$, we are
    looking at how the 26 letter predictors are derived from the
    previous layer's predictions. We see that most letters are
    negatively correlated (makes sense; only one will ever be 1.0)
    although there are some oddities (probably because it found some
    other, better correlates). Nodes on this new layer have dense
    references to these 26 predictions on the previous; this means
    that the bottom row kind of represents biases for each of the 26
    letters (what does an average 'e' look like?). I also included a
    dense region above that, but this appears to have simply evolved
    the same way as other $36^2$ chunks have (the rest are sparse).
    These chunks have a large amount of spatial similarity (suggesting
    that the sparse sampling would be adequate), with a meaning like
    ``if this area of the image is bright, then this pixel should
    be less bright.'' It is interesting that the pixels immediately
    next to the diagonal are almost always strongly negative
    (sharpening operation). ``Thank you for attending my TED talk.''
    ---\,Figure~\ref{fig:lastlayer}}
  \label{fig:lastlayer}
\end{figure}

@@

The actual error on the predicted SDFs is pretty low; for the
\makelowercase\ model it is around 31.3, which is like if 2.4\% of the
pixels were (completely) wrong, but the rest is exactly correct. In
reality, of course, the error is distributed throughout the pixels,
and some errors are a lot more important than others. Particularly,
near the threshold value, a pixel goes from from being considered ``in
the letter'' to ``outside'' with tiny changes in its value. Changes to
a pixel with a value near 0.0 or 1.0 usually doesn't affect the output
shape at all, in contrast. So one thing I did was map the loss
function (comparing expected pixel value to actual) to ``stretch out''
the region near the threshold, increasing the penalty (basically, the
derivative) in that region and decreasing it elsewhere. Looking at the
code again right now, I realize that I only applied this to the first
row of the SDF (\verb+idx < SDF_SIZE+ instead of
\verb+SDF_SIZE * SDF_SIZE+), so that was dumb AND MAKES ME {\bf
  ANGRY}. I will say in my defense that at least I felt disappointed
at the time that it didn't seem to make a difference! (The dark
wizard of superstitious fiddling nods sagely.

@@

\medskip
Now we're all set up to take an input shape (e.g.~from an existing
font), run the \makeuppercase\ or \makelowercase\ model(s) on it,
maybe multiple times, and trace the resulting SDF into a vector form
that can be used in a font. I did this on the canonical sans serif
font Helvetica~\cite{helvetica} and serif font Times New
Roman~\cite{timesnewroman}.

Helvetica means ``of Hell'', so making the font more uppercase give us
{\bf Heavenica} (Figure~\ref{fig:heavenica}), since Heaven is ``up''
from Hell. What's lower than Hell? Spezial Hell,\footnote{
  I first learned about Spezial Hell from a Rugen Br\"au beer that I
  drank in the Alps in Grindelwald, Switzerland ({\it la
    Conf\'ed\'eration H\'elvetique}).
  % Conf\oe{}deratio Helvetica
}
%
as in ``There's a Spezial Hell for the scalpers and cryptocurrency
environmental terrorists stockpiling GeForce 3000 series cards so that
I can {\em not} just get {\em one} darn card at a reasonable price for
my silly SIGBOVIK experiments.'' So the extra-lowercase version of
Helvetica is {\bf Spezial Hellvetica} (also Figure~\ref{fig:heavenica}).

@@

The UI also tells you how much your input resembles the various
existing letters \letterform{A}--\letterform{Z} and
\letterform{a}--\letterform{z} using the 26 bonus outputs that each
model predicts. For example, I learned that the ``Cool
S''~\cite{wikipediacools}:

@@

We can improve the results by searching for inputs with a ``perfect''
prediction (1.0) rather than making it as high as possible. These
results may not have been fully bonkers, but were at least downright
wacky. Since there appear to be a large variety of inputs that the
model judges as ``perfect'', the most appealing results from this
excursion involved scoring some additional properties of the
hallucinated inputs to discourage them from being so barmy. I
generated $8 \times 8$ bitmaps, plenty of pixels to make readable
letters on classic computers. Rather than allowing them to be
arbitrarily noisy, I also weakly optimized for (1) the number of
pixels set being close to half (2) minimal transitions between
on and off along each row and column. This produced shapes that
are basically letter-like, but weird (Figure~\ref{fig:perfect}).

@@

At each turn, the algorithm takes the board state that would
result from each legal move, and interprets it as an $8 \times 8$
bitmap. It renders that bitmap as an SDF and then runs the
\makelowercase\ model on it, and chooses the move that minimizes
the difference between its letter predictions and the
$[0, 0, \ldots 1, \ldots 0]$ vector corresponding to the letter
we are ``playing as.'' \deterministic \asymmetric

After tens of thousands of games each, it is clear that the letter-based
players are all bad at chess. Letters \letterform{E} and
\letterform{S} perform the worst (agreed on \letterform{S} being the
worst, thank you very much!), even worse than the ``No, I insist!''
strategy that tries to force its opponent to capture its pieces.
Letters \letterform{K} and \letterform{T} (eponymous! yeah!) perform
the best, but still worse than random. The numeric players like $\pi$
and $e$ are categorically better than the alphabetical ones, but this
is not surprising because chess is more of a mathematical game than a
linguistic one (Figure~\ref{fig:chess}).

@@

% figure? drop it?
{
  \tiny
\begin{tabular}{lrrrr}
name & elo & win & loss & draw \\
\hline  
worstfish & 395.95 & 10 & 44406 & 18584 \\
no i insist & 605.47 & 0 & 20405 & 42595 \\
letter f & 605.60 & 1341 & 21653 & 40006 \\
letter y & 606.00 & 1347 & 21688 & 39965 \\
letter b & 607.39 & 1705 & 21870 & 39425 \\
letter p & 607.61 & 1790 & 21963 & 39247 \\
letter l & 608.75 & 1370 & 21306 & 40324 \\
letter j & 610.86 & 1737 & 21525 & 39738 \\
letter u & 611.41 & 1363 & 20947 & 40690 \\
letter c & 612.38 & 1264 & 20787 & 40949 \\
huddle & 612.60 & 1172 & 20494 & 41334 \\
letter n & 612.81 & 1395 & 20802 & 40803 \\
letter w & 613.64 & 1342 & 20723 & 40935 \\
letter g & 613.72 & 1339 & 20660 & 41001 \\
letter v & 614.18 & 1390 & 20606 & 41004 \\
letter h & 615.04 & 1328 & 20452 & 41220 \\
letter r & 615.22 & 1826 & 20932 & 40242 \\
letter x & 615.70 & 1414 & 20457 & 41129 \\
letter m & 616.31 & 1774 & 20809 & 40417 \\
letter q & 618.85 & 1378 & 19993 & 41629 \\
letter o & 619.08 & 1502 & 20078 & 41420 \\
letter z & 620.09 & 1745 & 20275 & 40980 \\
letter d & 620.10 & 1730 & 20369 & 40901 \\
letter a & 620.57 & 1819 & 20212 & 40969 \\
letter i & 622.50 & 2169 & 20424 & 40407 \\
letter k & 622.90 & 1755 & 19878 & 41367 \\
letter t & 623.30 & 2237 & 20271 & 40492 \\
\ldots \\
random move & 655.90 & 7267 & 20983 & 34750 \\
\ldots \\
chessmaster.nes lv1 & 1014.74 & 37302 & 13992 & 11706 \\
\ldots \\
stockfish1m & 2831.66 & 60167 & 493 & 2340 \\
\end{tabular}
}

@@

People often stop me on the street to ask, Tom, Why did you spend so
much time and energy on this useless SIGBOVIK project? To which I say,
Ha! At least I am not wasting my time {\it reading} SIGBOVIK papers or
talking to strangers on the street! and run off. Although the main
purpose of SIGBOVIK is to confound bibliometrics with ambiguously
good-faith and high-quality research published in a clearly satirical
but superfically decorous venue of a traditionally esteemed university,
it is also possible for such work to have practical applications in
arts and spycraft. You simply need to give it some thought.

For example, it is well known to {\sf internet troIIs} and {\sf DOMAlN
  NAME PHlSHERS} that the uppercase \letterform{i} and loweracse
\letterform{L} are indistinguishable in many fonts, allowing for
various Tomfoolery.\footnote{I actually wrote ``troiis'' and ``domaln
  name phlshers'', hehe!} This can also be used for
steganography---hiding messages inside text---by replacing letters
with their alternates. Each replacement only encodes one bit of
information, however. With the generic ability to uppercase and
lowercase letterforms, we can exploit this ambiguity to generate
a large variety of letterforms that can be used like this.

@@

@@

% taller:
% adlnotwERTUZ
% 
% shorter:
% ceikmqrsyHQSWXY
% 
% Other applications: Steganography with llIIllIl

@@

We looked at what we get when we run \makelowercase\ on an existing
lowercase letter, making it lowerercase. Of course, we can run the
model again, and get an even lowererercase letter. The process can be
repeated indefinitely. As it turns out, lowercasing tends to make
letters smaller and smaller (makes sense) and they eventually just
turn to dust (Figure~\ref{fig:lowestcaseo}) and stay that way (makes
sense; ``dust to dust'').

@@

On the other hand, repeatedly uppercasing produces more interesting
results. Uppercasing usually increases the scale of a letter, but
this effect is limited by the finiteness of the SDF and the fact
that XXX

@@

\section{Acknowledgements}

For this project I used \verb+stb_truetype.h+ from the excellent stb
collection of single-file libraries~\cite{stb}. I did push its limits
somewhat and ``automatically discovered'' assertion failures and other
crashes (e.g.~during blackbox optimization on all 100k fonts), but it
saved a lot of time. This library only helps with reading the fonts
and generating SDFs. To generate TTFs, I had to write my own pipeline,
which generates FontForge's .SFD (a typo factory right there) files,
and then did the final export with FontForge~\cite{fontforge}. Thanks
Jason Reed for this suggestion, without which I would probably still
be developing my own TTF file writer and custom hinting engine. Finally,
I would like to thank the pseudonymous SIGBOVIK program committee for
overseeing the proceedings, and the \upsigbovik\ program committee
committee for overseeing {\em them}.

>>>>

% \usepackage{code}
% \usepackage{cite}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% IPA symbols. safe turns off overrides for like \! which I still want
\usepackage[safe]{tipa}
\usepackage{chessboard}
% \usepackage{chessfs}
% \usepackage{adjustbox}
\usepackage{supertabular}

@@

There's nothing wrong with this (despite the classical advice to use
shift to {\em reduce} conflict~\cite{aho1986compilers}). But the thing is:
What if I'm even MORE ANGRY THAN I WAS BEFORE? There are some standard
sorts of typographic emphasis, like I can be {\bf BOLD ANGRY} or
\textbf{\textit{\large BIG BOLD ITALIC UNDERLINE ANGRY}} or { \Large
  \textbf{\textit{\uuline{COMBINE A LOT OF THESE ANGERS}}}}, each with
its own nuances, depending on the cascading style sheet or LaTeX class
file. To be even more casual than lowercase, u can learn 2 write like
this, and {\scriptsize shrink away} and also \sout{cross out ur words
  in shame in advance of them even being read}, but there are few
other options for de-emphasis. Plus, when I'm FEELING PRETTY ANGRY,
TOM, how do I capitalize that already-capitalized T in order to show
the proper reverence for your humanity?

@@

@@

\begin{figure}[hb]
\centering
\includegraphics[width=0.9 \linewidth]{extracase}
\caption{%
  Probably someone already had this idea and did it before I was
  even born, thus taking the fun out of it for the rest of us, but
  here's a hand-made alphabet with ``upperercase'' and ``lowerercase''
  letters. You can download this TrueType font from \url{tom7.org/lowercase}.
  %
} \label{fig:manual}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.9 \linewidth]{training}
\caption{ Simultaneously training the two models. This example
  illustrates how a pair of letterforms \letterform{A} and
  \letterform{a} from the same font becomes four training examples.
  The pair straightforwardly generates an example
  \trainingexample{A}{a} for the \makelowercase\ queue, and an example
  \trainingexample{a}{A} for the \makeuppercase\ queue. Separately, we
  supply \letterform{a} to the \makelowercase\ model, simply to get
  the current output \letterform{\lowerlowera} (no model updates are
  performed). But this pair reversed becomes a training example
  \trainingexample{\lowerlowera}{a} for the \makeuppercase\ queue.
} \label{fig:cotraining}
\end{figure}

@@

\begin{figure}[ht!b]
\centering
  \includegraphics[width=0.99 \linewidth]{lastlayer}
  \caption{ The bottom-right corner of the weight matrix for the final
    layer of the network. This layer was added to the network after
    1.36 million rounds, initially as the identity matrix, and so can
    be thought of partly as a correction of the network's output prior
    to that round (though the remainder of the network continues to
    evolve). The x-axis is the output nodes, and the y-axis is the
    nodes of the previous layer. Note for example that the last 26
    columns look pretty different; these are the predictors for the 26
    letters, which occur in the output after the SDF pixels. Green
    means positive and red means negative, so if you are looking at
    this in a black-and-white printout, that may explain your current
    confusion. The exact diagonal is a strong green, close to 1.0,
    although over time these weights do diverge from the identity
    somewhat. In the bottom-right corner of size $26^2$, we are
    looking at how the 26 letter predictors are derived from the
    previous layer's predictions. We see that most letters are
    negatively correlated (makes sense; only one will ever be 1.0)
    although there are some oddities (probably because it found some
    other, better correlates). Nodes on this new layer have dense
    references to these 26 predictions on the previous; this means
    that the bottom row kind of represents biases for each of the 26
    letters (what does an average 'e' look like?). I also included a
    dense region above that, but this appears to have simply evolved
    the same way as other $36^2$ chunks have (the rest are sparse).
    These chunks have a large amount of spatial similarity (suggesting
    that the sparse sampling would be adequate), with a meaning like
    ``if this area of the image is bright, then this pixel should
    be less bright.'' It is interesting that the pixels immediately
    next to the diagonal are almost always strongly negative
    (sharpening operation). ``Thank you for attending my TED talk.''
    ---\,Figure~\ref{fig:lastlayer}}
  \label{fig:lastlayer}
\end{figure}

@@

The actual error on the predicted SDFs is pretty low; for the
\makelowercase\ model it is around 31.3, which is like if 2.4\% of the
pixels were (completely) wrong, but the rest is exactly correct. In
reality, of course, the error is distributed throughout the pixels,
and some errors are a lot more important than others. Particularly,
near the threshold value, a pixel goes from from being considered ``in
the letter'' to ``outside'' with tiny changes in its value. Changes to
a pixel with a value near 0.0 or 1.0 usually doesn't affect the output
shape at all, in contrast. So one thing I did was map the loss
function (comparing expected pixel value to actual) to ``stretch out''
the region near the threshold, increasing the penalty (basically, the
derivative) in that region and decreasing it elsewhere. Looking at the
code again right now, I realize that I only applied this to the first
row of the SDF (\verb+idx < SDF_SIZE+ instead of
\verb+SDF_SIZE * SDF_SIZE+), so that was dumb AND MAKES ME {\bf
  ANGRY}. I will say in my defense that at least I felt disappointed
at the time that it didn't seem to make a difference! (The dark
wizard of superstitious fiddling nods sagely.)

@@

\medskip
Now we're all set up to take an input shape (e.g.~from an existing
font), run the \makeuppercase\ or \makelowercase\ model(s) on it,
maybe multiple times, and trace the resulting SDF into a vector form
that can be used in a font. I did this on the canonical sans serif
font Helvetica~\cite{helvetica} and serif font Times New
Roman~\cite{timesnewroman}. Each font is produced by a symmetric
process; for example, to make a font ``more lowercase,'' I take the
input font's lowercase alphabet and run \makelowercase\ on it (this
becomes the output font's lowercase letters), and then run
\makeuppercase\ on {\em those} to produce the output font's uppercase
letters. These letters are usually recognizable as the ``normal''
lowercase letters even though they've been through both neural
networks. Before tracing, I do some automatic gamma adjustment of the
SDFs (e.g.~at least 5\% of the pixels should be above the threshold),
as the unadjusted letters seemed a bit too light. These fonts can
also be downloaded from \url{http://tom7.org/lowercase/} for your
corporate PowerPoint needs.

Helvetica means ``of Hell'', so making the font more uppercase give us
{\bf Heavenica} (Figure~\ref{fig:heavenica}), since Heaven is ``up''
from Hell. What's lower than Hell? Spezial Hell,\footnote{
  I first learned about Spezial Hell from a Rugen Br\"au beer that I
  drank in the Alps in Grindelwald, Switzerland ({\it la
    Conf\'ed\'eration H\'elvetique}).
  % Conf\oe{}deratio Helvetica
}
%
as in ``There's a Spezial Hell for the scalpers and cryptocurrency
environmental terrorists stockpiling GeForce 3000 series GPUs so that
I can {\em not} just get {\em one} darn card at a reasonable price for
my important SIGBOVIK experiments.'' So the extra-lowercase version of
Helvetica is {\bf Spezial Hellvetica} (also Figure~\ref{fig:heavenica}).

@@

The UI also tells you how much your input resembles the various
existing letters \letterform{A}--\letterform{Z} and
\letterform{a}--\letterform{z} using the 26 bonus outputs that each
model predicts. For example, I learned that the
``Cool~S''~\cite{wikipediacools}:

@@

We can improve the results by searching for inputs with a ``perfect''
prediction (1.0) rather than making it as high as possible. These
results may not have been fully bonkers, but were at least downright
wacky. Since there appear to be a large variety of inputs that the
model judges as ``perfect'', the most appealing results from this
excursion involved scoring some additional properties of the
hallucinated inputs to discourage them from being so barmy. I
generated $8 \times 8$ bitmaps, plenty of pixels to make readable
letters on classic computers. Rather than allowing them to be
arbitrarily noisy, I also weakly optimized for (1) the number of
pixels set being close to half and (2) minimal transitions between on
and off along each row and column. This produced shapes that are
basically letter-like, but weird (Figure~\ref{fig:perfect}).

@@

At each turn, the algorithm takes the board state that would
result from each legal move, and interprets it as an $8 \times 8$
bitmap. It renders that bitmap as an SDF and then runs the
\makelowercase\ model on it, and chooses the move that minimizes
the difference between its letter predictions and the
$[0, 0, \ldots 1, \ldots 0]$ vector selecting the letter
we are ``playing as.'' \deterministic \asymmetric

After tens of thousands of games each, it is clear that the
letter-based players are all bad at chess. Letters \letterform{E} and
\letterform{S} perform the worst (agreed on \letterform{S} being the
worst, thank you very much!), even worse than the ``No, I insist!''
strategy that tries to force its opponent to capture its pieces. The
letter \letterform{T} (eponymous! yeah!) performs best, but still
worse than random. The numeric players like $\pi$ and $e$ are
categorically better than the alphabetical ones, but this is not
surprising because chess is more of a mathematical game than a
linguistic one (Figure~\ref{fig:chess}).

@@

% figure? drop it?

The abbreviated tournament results:
{
\scriptsize
\begin{center}

\tablefirsthead{%
  {\bf name} & {\bf elo} & {\bf wins} & {\bf losses} & {\bf draws}   \\  
  \hline
  }
\tablehead{%
  \ldots & {\bf elo} & {\bf wins} & {\bf losses} & {\bf draws}   \\    
  \hline
}
\tabletail{%
  \multicolumn{5}{c}{\ldots} \\
  % \hline
}
\tablelasttail{%
  % nothing
}
% \begin{tabular}{lrrrr}
\begin{supertabular}{lrrrr}
% name & elo & wins & losses & draws \\
% \hline  
worstfish & 395.95 & 10 & 44406 & 18584 \\
\ldots \\
letter e &  602.27 &  1717 & 22566 & 38717 \\
letter s &  602.92 &  1310 & 22020 & 39670 \\
no i insist & 605.47 & 0 & 20405 & 42595 \\
letter f & 605.60 & 1341 & 21653 & 40006 \\
letter y & 606.00 & 1347 & 21688 & 39965 \\
letter b & 607.39 & 1705 & 21870 & 39425 \\
letter p & 607.61 & 1790 & 21963 & 39247 \\
letter l & 608.75 & 1370 & 21306 & 40324 \\
letter j & 610.86 & 1737 & 21525 & 39738 \\
letter u & 611.41 & 1363 & 20947 & 40690 \\
letter c & 612.38 & 1264 & 20787 & 40949 \\
huddle & 612.60 & 1172 & 20494 & 41334 \\
letter n & 612.81 & 1395 & 20802 & 40803 \\
letter w & 613.64 & 1342 & 20723 & 40935 \\
letter g & 613.72 & 1339 & 20660 & 41001 \\
letter v & 614.18 & 1390 & 20606 & 41004 \\
letter h & 615.04 & 1328 & 20452 & 41220 \\
letter r & 615.22 & 1826 & 20932 & 40242 \\
letter x & 615.70 & 1414 & 20457 & 41129 \\
letter m & 616.31 & 1774 & 20809 & 40417 \\
letter q & 618.85 & 1378 & 19993 & 41629 \\
letter o & 619.08 & 1502 & 20078 & 41420 \\
letter z & 620.09 & 1745 & 20275 & 40980 \\
letter d & 620.10 & 1730 & 20369 & 40901 \\
letter a & 620.57 & 1819 & 20212 & 40969 \\
letter i & 622.50 & 2169 & 20424 & 40407 \\
letter k & 622.90 & 1755 & 19878 & 41367 \\
letter t & 623.30 & 2237 & 20271 & 40492 \\
\ldots \\
random move & 655.90 & 7267 & 20983 & 34750 \\
\ldots \\
chessmaster.nes lv1 & 1014.74 & 37302 & 13992 & 11706 \\
\ldots \\
stockfish1m & 2831.66 & 60167 & 493 & 2340 \\
\end{supertabular}
% \end{tabular}
\end{center}
}

@@

People often stop me on the street to ask, Tom, Why did you spend so
much time and energy on this useless SIGBOVIK project? To which I say,
Ha! At least I am not wasting my time {\it reading} SIGBOVIK papers or
talking to strangers on the street! and run off. Although the main
purpose of SIGBOVIK is to confound bibliometrics with ambiguously
good-faith and high-quality research published in a clearly satirical
but superfically decorous venue associated with a traditionally
esteemed university, it is also possible for such work to have
practical applications in arts and spycraft. You simply need to give
it some thought.

For example, it is well known to {\sf internet troIIs} and {\sf DOMAlN
  NAME PHlSHERS} that the uppercase \letterform{i} and loweracse
\letterform{L} are indistinguishable in many fonts, allowing for
various Tomfoolery.\footnote{I actually wrote ``troiis'' and ``domaln
  name phlshers'', hehe!} This can also be used for
steganography---hiding messages inside text without the use of em
dashes---by selectively replacing letters with their alternates. Each
replacement only encodes one bit of information, however. With the
generic ability to uppercase and lowercase letterforms, we can exploit
this ambiguity to generate a large variety of letterforms that can be
used like this.

@@

% letters in figure:
% adlnotwERTUZ

@@

@@

We looked at what we get when we run \makelowercase\ on an existing
lowercase letter, making it lowerercase. Of course, we can run the
model again, and get an even lowererercase letter. The process can be
repeated indefinitely. As it turns out, lowercasing tends to make
letters smaller and smaller (makes sense) and they eventually just
turn to dust (Figure~\ref{fig:lowestcaseo}) and stay that way (makes
sense; ``dust to dust''~\cite{commonprayerburial}).

@@

On the other hand, repeatedly uppercasing produces more interesting
results. Uppercasing usually increases the scale of a letter, but this
effect is limited by the finiteness of the SDF and the fact that the
outer edges are very unlikely to have high values. Moreover, although
uppercase letters are large, they also have a lot of internal space.
So iterations do not simply grow in size or fill the space, but
repeatedly grow and deteriorate like an organism in the Game of
Life~\cite{games1970fantastic}. Animating the SDF under iterations of
the model is reminiscent of a flickering candle, recalling another
well-known approach to increasing the emphasis of text,
\url{flamingtext.com}. Is this eternal flame itself the uppestcase
letter? Alas, such an effect is not possible in print (paper is too
flammable).

Iterating the \makeuppercase\ model with 32-bit floats does not form
any cycles in 25,000,000 iterations, nor does it if the intermediates
are quantized to 8-bit ints. This is curious because under visual
observation the sequence does appear periodic, and in fact seems to be
the same characteristic loop reached by all letters. Presumably there
are some oscillations with long, relatively prime periods or even some
pixels that are monotonically growing or shrinking, but very slowly.
However, if I render the SDF as a 1-bit bitmap at $2\times$ scale, I
get a perfect cycle of 132 frames, which appears to be two passes
through the main characteristic loop, but slightly different each
time. This is not a strange property for a letter to have; for example
a typical \letterform{B} has two copies of the same basic idea in it.

.... XXX more here ...

@@

\section{Conclusion}
... XXX ...
This fixed point
suggests that the lowestcase letter is already on your keyboard,
but it is not a point, it is just \letterform{:}.

\paragraph{Futura Work}
I think I pretty much beat this one to death, honestly.

\paragraph{Acknowledgements}
For this project I used \verb+stb_truetype.h+ from the excellent stb
collection of single-file libraries~\cite{stb}. I did push its limits
somewhat and ``automatically discovered'' assertion failures and other
crashes (e.g.~during blackbox optimization on all 100k fonts), but it
saved a lot of time. This library only helps with reading the fonts
and generating SDFs. To generate TTFs, I had to write my own pipeline,
which generates FontForge's .SFD (a typo factory right there) files,
and then did the final export with FontForge~\cite{fontforge}. Thanks
Jason Reed for this suggestion, without which I would probably still
be developing my own TTF file writer and custom hinting engine. Finally,
I would like to thank the pseudonymous SIGBOVIK program committee for
overseeing the proceedings, and the \upsigbovik\ program committee
committee for overseeing {\em them}.
